---
title: "Project 1"
output: html_document
date: "2023-10-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```










```{r}
# tidy life expectancy
merged_data_2020 = read.csv('merged_data_2020.csv')


merged_data_2020 = 
  merged_data_2020 |>
  select(Country, Continent, Hemisphere, everything())





# summary stats
# world mean GDP in $
merged_data_2020 |>
  summarize(mean_gdp = mean(GDP_Value, na.rm = TRUE))

# std dev of world mean GDP in $
merged_data_2020 |>
  summarize(sd_gdp = sd(GDP_Value, na.rm = TRUE))


# mean life expectancy at birth in years by continent
merged_data_2020 |>
  group_by(Continent) |>
  summarize(mean_life_expectancy = mean(Life_Expectancy_at_Birth, na.rm = TRUE)) |>
  arrange(desc(mean_life_expectancy))

 


```

```{r}
library(ggplot2)
# histogram by continent
merged_data_2020 |>
  ggplot(aes(x = Continent)) +
  geom_histogram(stat = 'count') + 
  labs(x = 'Continent', y = 'Count of Countries', title = 'Histogram of Countries per Continent') + 
  theme_minimal() + 
  scale_y_continuous(limits = c(0, 60))




# histogram by hemisphere
merged_data_2020 |>
  ggplot(aes(x = Hemisphere)) +
  geom_histogram(stat = 'count') + 
  labs(x = 'Hemisphere', y = 'Count of Countries', title = 'Histogram of Countries per Hemisphere') + 
  theme_minimal() + 
  scale_y_continuous(limits = c(0, 160))




```




```{r}
# Two Variable Plots
# growth rate vs life expectancy
merged_data_2020 |>
  ggplot(aes(x = Growth.Rate, y = Life_Expectancy_at_Birth)) + 
  geom_point() + 
  labs(x = 'Growth Rate (%)', y = 'Life Expectancy at Birth (years)', title = 'Scatter Plot of Growth Rate vs Life Expectancy') + 
  theme_minimal() + 
  scale_x_continuous(n.breaks = 7)




```


```{r}
# 2 variable 
# human development groups vs GDP
merged_data_2020 |>
  group_by(Human.Development.Groups) |>
  summarize(avg_gdp = mean(GDP_Value, na.rm = TRUE)) |>
  ggplot(aes(x = factor(Human.Development.Groups, levels = c('Low', 'Medium', 'High', 'Very High')), y = avg_gdp)) + 
  geom_bar(stat = 'summary') + 
  labs(x = 'Human Development Groups', y = 'Average GDP ($)', title = 'Bar Graph of Average GDP by Human Development Groups') + 
  theme_minimal() + 
  scale_y_continuous(n.breaks = 7)
```








```


